Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-05-31 doc/network.tex
tweak

Population-level method tries to estimate the effective generation-interval distribution by reverting the inverse exponential weighting in the observed generation-interval distribution:
\begin{equation}
\tsub{g}{eff}(\tau) \propto c_t(\tau) \exp(r\tau).
\end{equation}
In order to do so, we first fit a gamma distribution to the observed generation-interval distributions; specifically, we esimate the mean and the shape parameter of a gamma distribution by maximum likelihood.
\swp{Come back}

@@

The entire data likelihood of the individual-level method can be written as:
\begin{equation}
\begin{aligned}
&\mathcal{L}(\RR, \theta\,|\, \mathbf{s}, \mathbf{t}, \mathbf{n}, \tsub{t}{censor})\\
&\quad=\prod_{j=1}^N \left(\RR^{n_j} \exp \big(- \RR \tsub{G}{eff}(\tsub{t}{censor} - t_j ;\theta) \big) \prod_{i=1}^{n_j} \tsub{g}{eff}(s_{i, j}; \theta) \right),
\end{aligned}
\end{equation}
where $N$ is the total number of infected individuals.
Here, we estimate parameters $\RR$ and $\theta$ by maximum likelihood.
Since the estimate of $\RR$ is sensitive to under-reporting, we use the estimated generation-interval distribution to infer the effective reproductive number from the estimated growth rate.
\cite{forsberg2008likelihood} proposed a related approach based on discretized incidence reports.

\dc{that reminds me of Cauchemez's 2006 AJE paper: ``Estimating in Real Time the Efficacy of Measures to Control Emerging Communicable''. Not sure if it's really relevant, but just in case...  }

>>>>

Population-level method tries to estimate the effective generation-interval distribution by reverting the inverse exponential weighting in the observed generation-interval distribution:
\begin{equation}
\tsub{g}{eff}(\tau) \propto c_t(\tau) \exp(r\tau).
\end{equation}
In order to do so, we first fit a gamma distribution to the observed generation-interval distributions; specifically, we esimatemean $\bar G$ and shape $\alpha$ of a gamma distribution by maximum likelihood.
Then, the effective generation-interval distribution also follows a gamma distribution with mean $\alpha/(\alpha/\bar G - r)$ and shape $\alpha$.

@@

The entire data likelihood of the nonhomogeneous poisson process can be written as:
\begin{equation}
\begin{aligned}
&\mathcal{L}(\RR, \theta\,|\, \mathbf{s}, \mathbf{t}, \mathbf{n}, \tsub{t}{censor})\\
&\quad=\prod_{j=1}^N \left(\RR^{n_j} \exp \big(- \RR \tsub{G}{eff}(\tsub{t}{censor} - t_j ;\theta) \big) \prod_{i=1}^{n_j} \tsub{g}{eff}(s_{i, j}; \theta) \right),
\end{aligned}
\end{equation}
where $N$ is the total number of infected individuals.
Here, we estimate parameters $\RR$ and $\theta$ by maximum likelihood.
Since the estimate of $\RR$ is sensitive to under-reporting, we use the estimated generation-interval distribution to infer the effective reproductive number from the estimated growth rate.
\cite{forsberg2008likelihood} proposed a related approach based on discretized incidence reports.
